---
title: "PW4 - Logistic Regression"
author: "Sandrine NEANG"
date: "15/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<b>Social Networks Ads</b><br>

<b>Question 1 : Import the Social_Network_Ads dataset into R.</b><br>
```{r question1, echo=TRUE}
Social_Network_Ads <- read.csv("~/Scolarité/ESILV/M1 A4/S7/Machine Learning/TDs/Datas/Social_Network_Ads.csv")
```
<br>

<b>Question 2 : Describe the dataset.</b><br>
```{r question2, echo=TRUE}
str(Social_Network_Ads)
summary(Social_Network_Ads)
```

<b>Question 3 : Split the dataset ramdomly into training set and test set with caTools packages.</b><br>
```{r question3, echo=TRUE}
library(caTools)
set.seed(123)
sample <- sample.split(Social_Network_Ads, SplitRatio = 0.75)
train <- subset(Social_Network_Ads, sample==TRUE)
test <- subset(Social_Network_Ads, sample==FALSE)
```

<b>Question 4 : Scale the input variables (Age and EstimatedSalary) in both training set and test set.</b><br>
```{r question4, echo=TRUE}
train$Age = (train$Age - mean(train$Age))/sd(train$Age)
train$EstimatedSalary = (train$EstimatedSalary - mean(train$EstimatedSalary))/sd(train$EstimatedSalary)

test$Age = (test$Age - mean(test$Age))/sd(test$Age)
test$EstimatedSalary = (test$EstimatedSalary - mean(test$EstimatedSalary))/sd(test$EstimatedSalary)

head(train)
head(test)
```

<b>Question 5 : Fit a simple logistic regression model of Purchased in function of Age.</b><br>
```{r question5, echo=TRUE}
model1 = glm(formula = Purchased ~ Age, data=train, family=binomial)
```

<b>Question 6 : In the argument family of the function glm, we chose binomial. Why?</b></br>
We chose binomial family because the Purchased variable is a categorical variable which using numerical predictors.

<b>Question 7 : What is the equation of the obtained model in the question 5?</b><br>
```{r question7, echo=TRUE}
summary(model1)
model1$coefficients
```
The equation is : Purchased = 2.0252322 *Age - 0.8963671<br><br>

<b>Question 8 : Is the feature Age significant?</b><br>
The p-value of Age variable is significantly smaller than 0.05 so Age is significant.<br><br>

<b>Question 9 : What is the value of AIC of the model ?</b><br>
```{r question9, echo=TRUE}
AIC(model1)
```

<b>Question 10 : Plot Purchased in function of Age and add the curve of the obtained logistic regression model.</b><br>
```{r question10, echo=TRUE}
library(ggplot2)

ggplot(train, aes(x=train$Age, y=train$Purchased)) + geom_point() +
      stat_smooth(method="glm", color="green", se=FALSE,
                method.args = list(family=binomial))
```

<b>Question 11 : Fit a logistic regression model of Purchasing the product in function of the Age of the user and its Salary.</b><br>
```{r question11, echo=TRUE}
model2 = glm(formula = Purchased ~ Age + EstimatedSalary, data=train, family=binomial)
summary(model2)
```

<b>Question 12 : Are the predictors significant?</b><br>
The p-value of the Age and EstimatedSalary variables are below 0.05 which means that the null hypothesis can be rejected and our results likely did not happen by chance
therefore the predictors are significant in this logistic regression model.
<br><br>

<b>Question 13 : Did the model get better by adding the EstimatedSalary?</b><br>
By adding the EstimatedSalary, the model get better.<br><br>


<b>Question 14 : On the test set, predict the probability of purchasing the product by the users using the obtained model.</b><br>
```{r question14, echo=TRUE}
Predicted_data <- predict(model2, test, typ = "response")
```

<b>Question 15 : We predicted the probability that the user will purchase the product. Now in order to compare our results with the real answers, transform the predicted values to 0 or 1 (1 if > 0.5).</b><br>
```{r question15, echo=TRUE}
Predicted_data <- ifelse(Predicted_data > 0.5, 1, 0 )
Predicted_data
```

<b>Question 16 : To evaluate the predictions, compute the confusion matrix. What do you obtain ?</b><br>
```{r question16, echo=TRUE}
library(caret)
matrix <- confusionMatrix(factor(Predicted_data), factor(test$Purchased))
matrix
```

<b>Question 17 : Calculate the accuracy, specificity, sensitivity and the precision of the model.</b><br>
```{r question17, echo=TRUE}
accuracy = matrix$overall['Accuracy']
accuracy
specificity = matrix$byClass['Specificity']
specificity
sensitivity = matrix$byClass['Sensitivity']
sensitivity
precision = matrix$byClass['Pos Pred Value'] 
precision
```

<b>Question 18 : Plot the ROC curve and calculate AUC value.</b><br>
```{r question18, echo=TRUE}
library(ROCR)

pred <- prediction(Predicted_data, test$Purchased)

roc=performance(pred,"tpr","fpr")
plot(roc ,main ="ROC curve - Logistic Regression ")


auc_ROCR <- performance(pred, measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
auc_ROCR
```

<b>Question 19 : Compare the AUC of the two models you fitted (one with only age and one with age and estimated salary) and plot their ROC curves in the same figure.</b><br>
```{r question19, echo=TRUE}
Predicted2 <- predict(model1, test,typ = "response")
pred_2 <- prediction(Predicted2, test$Purchased)

auc_2<- performance(pred_2, measure = "auc")
auc_2 <- auc_2@y.values[[1]]
auc_2

roc_2=performance(pred_2,"tpr","fpr")
plot(roc_2 ,main ="ROC curve - Logistic Regression ")

plot(roc, colorize = TRUE)
plot(roc_2, add = TRUE, colorize = TRUE)
```




